Estimating Operator Intent\\ (Draft 2.0)

### 2015-09-20 intent.tex
split into 3 papers

\title{A Survey of Exoskeleton Control (Draft 1.1)}

@@

\section{What is this paper about?}

This paper surveys state of the art exoskeleton technology with an
emphasis on control. The focus is on exoskeleton technologies that allow a
highly trained and top percentile athletic 
operator to carry a payload that weighs approximately the same amount
as the operator. We envisage these types of exoskeletons to be useful
in carrying protective and safety equipment for SWAT teams, police,
firefighters, and soldiers.

We expect each exoskeleton controller
to be used by and optimized for a single operator.
A substantial investment in capturing the operators normal behavior,
operator training and learning, and controller customization can be made.

We do not survey exoskeleton technology for human arms. 
We treat the torso and helmet of the exoskeleton as the payload,
and focus on a lower body exoskeleton to support any payloads that
are on the torso or head.
We do not cover global energy storage such as batteries or liquid fuel.

\section{What would it be like for the operator?}

\subsection{``Invisible'' exoskeleton}

Can we build an exoskeleton allows the operator to behave naturally
and exerts little or no force on the operator?

This could be achieved through active control using feedback of the
operator-exoskeleton forces. Additional gravity, friction, actuator, and
inertial compensation (inverse dynamics) can improve performance.

\subsection{``Natural'' exoskeleton}

Can we build an exoskeleton allows the operator to behave naturally?
The operator can feel the exoskeleton, but can overpower it when necessary.

Bypass valves and clutches allow the operator to physically ``take over''
and push the relatively lightweight leg mounted exoskeleton around.

``Passive'' degrees of freedom can have physical or ``virtual'' springs
applied to assist in support (all abduct/adduct DOF, rotational DOF).
These aids can be clutched out or turned off when necessary.
``Virtual'' sprintgs can be modulated.

\subsection{``Symbiotic'' exoskeleton}

Can we build an exoskeleton that allows the operator/exoskeleton team
to be effective? The operator needs to learn to ``fly'' the suit,
just as operators learn to operate parachutes,
wing suits, diving equipment, high and low temperature protective
suits, firefighting equipment, and high altitude flight suits.

The exoskeleton can adapt its behavior to the task.
For example, for jumping, the exoskeleton can operate like a pogo 
stick \url{https://www.youtube.com/watch?v=lbp41vWP4o4}, 
and for running it could act like jumping 
stilts \url{https://www.youtube.com/watch?v=9ZOd7yEyhwI},

\subsection{Physical Constraints}

1) There is a weight distribution issue. Ideally the exoskeleton should
have a center of mass at the same location as the operator. Otherwise,
operator movements can be degraded and substantial forces occur between
the operator and the exoskeleton.

This constraint can be reduced with the use of propulsive forces other
than foot contacts: rockets (solid or liquid fuel), jets, exploding armor,
ducted fan air flow, ...

A later white paper will discuss the use of propulsive forces other
than foot contacts, as well as additional actuation at the contacts (air
bags, springs, other ways to rapidly apply stored energy) for
operator safety (dodging threats, ...) and high performance (jumping, ...).

\section{Recognizing Operator Intent}

Survey current operator/exoskeleton interfaces.

The further in the future the control system can predict what the operator
is going to do, the easier it is to control the exoskeleton. One way
to think about predicting operator intent is to imagine you are in
a hand-to-hand fight and you want to predict your opponent's next move.
You can also think about playing poker. What ``tells'' does the opponent
have?
You are allowed to instrument your opponent. What sensors would you use?
How would you interpret the sensors?

@@

\subsection{Intent time scales}

@@

\subsection{Signal flow analysis}

@@

\subsubsection{What signals are there?}

@@

A future white paper will discuss how intent prediction could be 
made easier with superhuman perception.

A future white paper will discuss how intent prediction could be 
made easier with implanted markers and devices, as well as continuous
operator imaging such as ultrasound.

\subsubsection{Signal timing?}

So for inferring the ``intent'' of a human operator of an
exoskeleton,
I would like to put a timeline together of events.
I guessed some timing, which I hope you will correct.

When I say leg EMG or leg muscle I mean
gastrocneumius or vastus lateralis (``fast'' muscles).

What are delays between:
\begin{verbatim}
visual event to leg EMG start
auditory event to leg EMG start
vestibular event to leg EMG start
foot tactile event to leg EMG start
EMG start to 10% leg muscle
EMG start to peak leg muscle
\end{verbatim}

What are your guesses for motorneuron activation to leg EMG activation?

{\b From Hartmut Geyer:}

@@

\subsubsection{What can be measured or estimated?}

Predictive models of human motion, a major topic of robot learning
from observation/demonstration/imitiation and computer animation,
will be the subject of a future white
paper~\cite{IEEE06913830,Bagnell}

@@

{\bf EMG}
Electrical activity was measured using surface electrodes,
high-pass filtered at 20 Hz, rectified, low-pass filtered at 6 Hz,
offset by a small value of -0.008, and amplified by a gain of 283
to obtain the desired torque.~\cite{IEEE07139980}

muscle moves -> surface electrodes come off (especially for large fast
movements, the worst time).
perspration changes measurement. needs to be calibrated for each application.

@@

\subsubsection{How can the operator communicate with the exoskeleton}

@@

\subsection{Operator load and fatigue}

>>>>

\title{Estimating Operator Intent\\
(Draft 2.0)}

@@

\section{Scope: What is this paper about?}

This paper surveys state of the art technology to recognize
and predict exoskeleton
operator intent.
A companion paper surveys current operator/exoskeleton interfaces~\cite{}.

The focus is on intent estimation that allows a
highly trained and top percentile athletic 
operator to carry a payload that weighs approximately the same amount
as the operator. We envisage these types of exoskeletons to be useful
in carrying protective and safety equipment for SWAT teams, police,
firefighters, and soldiers.

We focus this survey on intent estimation for lower body tasks (standing, walking,
running, jumping, kicking, dodging, ...).
We do not survey intent estimation for manipulation.

A future white paper will discuss how intent estimation could be 
made easier with superhuman perception (sensing beyond human capabilities
or modalities).

A future white paper will discuss how intent estimation could be 
made easier with implanted markers and devices, as well as continuous
operator imaging such as ultrasound.

Predictive models of human motion, a major topic of robot learning from\\
observation/demonstration/imitation and computer animation,
will be the subject of a future white
paper~\cite{IEEE06913830,Bagnell}

\section{Introduction: Our Point of View}

{\bf What is intent estimation?}
The further in the future the control system can predict what the operator
is going to do, the easier it is to control the exoskeleton. One way
to think about estimating operator intent is to imagine you are in
a hand-to-hand fight and you want to predict your opponent's next move.
You can also think about playing poker. What ``tells'' does the opponent
have?
You are allowed to instrument your opponent. What sensors would you use?
How would you interpret the sensors?

@@

{\bf Individualization of Intent Estimation:}
We expect each exoskeleton controller
to be used by and optimized for a single operator.
A substantial investment in capturing the operators normal behavior,
operator training and learning, and controller customization can be made.

\section{Intent time scales}

@@

\section{What signals are there?}

@@

@@

\section{Signal timing?}

{\it So for inferring the ``intent'' of a human operator of an
exoskeleton,
I would like to put a timeline together of events.
When I say leg EMG or leg muscle I mean
gastrocneumius or vastus lateralis (``fast'' muscles).
What are delays between:
\begin{verbatim}
visual event to leg EMG start
auditory event to leg EMG start
vestibular event to leg EMG start
foot tactile event to leg EMG start
EMG start to 10% leg muscle
EMG start to peak leg muscle
\end{verbatim}
What are your guesses for motorneuron activation to leg EMG activation?
}

{\bf From Hartmut Geyer:}

@@

\section{What can be measured or estimated?}

{\bf EMG:}
Electrical activity was measured using surface electrodes,
high-pass filtered at 20 Hz, rectified, low-pass filtered at 6 Hz,
offset by a small value of -0.008, and amplified by a gain of 283
to obtain the desired torque.~\cite{IEEE07139980}

Muscle movement leads to surface electrodes losing good contact, fully detaching, or
even ripping off (especially for large fast
movements, the worst time).
Perspiration changes measurement. EMG measurement
needs to be calibrated for each electrode application and periodically during use.

@@

@@

\section{How can the operator communicate with the exoskeleton}

@@

\section{Operator load and fatigue}
